---
title: "INGARCH Modeling"
author: "Adam Kiehl"
date: "2/12/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(acp) # INAR model
library(tscount)

set.seed(499)
```

## Introduction

The univariate INGARCH(p, q) model fit here using the `tsglm` function in the 
`tscount` package is specified according to Davis, et al. like:
$$
\begin{aligned}
X_t|F_{t-1} &\sim Poisson(\lambda_t) \\
\lambda_t &= d + \sum_{j=1}^p b_{j} X_{t-j} + \sum_{j=1}^q a_j \lambda_{t-j}
\end{aligned}
$$
The marginal mean of the response poisson distribution is described in a
manner resembling a classical ARMA model, with the `p` parameter defining the 
extent of the autoregressive component and the `q` parameter defining the moving
average component. 

## Setup

The binned data frame is read and feature names are defined:
```{r}
binned_df <- read.csv('../Data/binned_df.csv')

resp_names <- c('CA1.1', 'CA1.2', 'CA1.3', 'CA1.4', 'CA1.5', 'CA1.6', 'CA1.7', 
                'CA1.8', 'CA1.9', 'CA1.10', 'CA1.11', 'CA1.12', 'CA1.13', 
                'CA1.14', 'CA1.15')
pred_names <- c('CA3.1', 'CA3.2', 'CA3.3', 'CA3.4', 'CA3.5', 'CA3.6', 'CA3.7', 
                'CA3.8', 'CA3.9', 'CA3.10', 'CA3.11', 'CA3.12', 'CA3.13', 
                'CA3.14', 'CA3.15', 'CA3.16', 'CA3.17', 'CA3.18', 'CA3.19', 
                'CA3.20', 'CA3.21')
```

A function for subsetting and cleaning data frames for modeling:
```{r}
gen_df <- function(train_trials, valid_trials, pred, resp) {
  # generate training set
  train_df <- binned_df %>%
    filter(trial %in% as.vector(train_trials)) %>%
    select(c(all_of(resp), all_of(pred), trial)) %>%
    data.frame()
  
  # generate validation set
  valid_df <- binned_df %>%
    filter(trial %in% as.vector(valid_trials)) %>%
    select(c(all_of(pred), trial)) %>%
    data.frame()
  
  # values to compare predicted values with
  valid_pred <- binned_df %>%
    filter(trial %in% as.vector(valid_trials)) %>%
    select(all_of(resp))
  
  # handle empty vectors
  zeros <- c()
  zeros <- c(zeros, which(colSums(train_df) == 0))
  zeros <- c(zeros, which(colSums(valid_df) == 0))
  if (length(zeros) > 0) {
    train_df <- train_df[, -zeros]
    valid_df <- valid_df[, -zeros]
  }
  
  # return data frames
  return(list('train_df' = train_df, 
              'valid_df' = valid_df, 
              'valid_pred' = valid_pred))
}
```

A function for cross-validation given the parameters of a model specification:
```{r}
cross_valid <- function(pred_list, resp_list, p, q) {
  # define empty vector to store MSE values
  mses <- rep(NA, 2)
  
  for (i in 1:2) {
    # select trials for training and validation (PICK ONE)
    # blocking cross-validation method is progressive
    trials <- list((10 * (i - 1) + 1):(10 * i - 2), (10 * i - 1):(10 * i))
    # time series split cross-validation method is inclusive
    # trials <- list(1:(10 * i - 2), (10 * i - 1):(10 * i))
    
    # collect data frames
    dfs <- gen_df(trials[[1]], trials[[2]], pred_list, resp_list)
    train_df <- dfs[['train_df']]
    valid_df <- dfs[['valid_df']][1:100, ]
    valid_pred <- dfs[['valid_pred']][1:100, ]
    remove(dfs)
    
    tryCatch(
      {
        # fit INGARCH model
        model <- tsglm(as.ts(train_df[, 1]), xreg = as.matrix(train_df[, -1]), 
                       model = list(past_obs = 1:p, past_mean = q), 
                       link = 'log', distr = 'poisson')
      }, 
      error = function(err) {
        print(paste('Model fitting error on (', p, ', ', q, '): ', err, 
                    sep = ''))
      }
    )
    
    tryCatch(
      {
        # calculate MSE from predicted values
        pred <- predict(model, n.ahead = nrow(valid_df), newxreg = valid_df)
        mse <- mean((pred[['pred']] - valid_pred)^2, na.rm = TRUE)
        
        mses[i] <- round(mse, 4)
      }, 
      error = function(err) {
        print(paste('Model evaluation error on (', p, ', ', q, '): ', err, 
                    sep = ''))
      }
      
    )
  }
  
  return(mean(mses, na.rm = TRUE))
}
```

# Variable Selection

Perform variable selection on the INGARCH model using a forward selection 
technique based on the effect sizes from the full model. 
```{r, message=FALSE, error=FALSE}
# define empty list to store best variable subsets
var_subset = list()

for (resp in resp_names) {
  # collect training data frame
  dfs <- gen_df(1:3, 21, pred_names, resp)
  train_df <- dfs[['train_df']]
  remove(dfs)

  tryCatch(
    {
      # fit full model
      model <- tsglm(as.ts(train_df[, 1]), xreg = as.matrix(train_df[, -1]), 
               model = list(past_obs = 1, past_mean = 2), 
               link = 'log', distr = 'poisson')
      
      # extract estimated coefficients from fitted model
      temp <- summary(model)
      est <- temp$coefficients$Estimate[-c(1:3, 25)]
      
      full_model_select <- data.frame(var = pred_names, coefs = est) %>% 
        arrange(desc(coefs)) %>%
        top_n(10) %>%
        select(var) %>%
        unlist()
    },
    error = function(err) {
      print(paste('Error:', err))
    }
  )
  
  # define empty vector to store CV MSE results for forward selection
  mses <- rep(NA, 10)
  
  # perform forward selection with chosen variables
  for (i in 1:length(full_model_select)) {
    preds <- full_model_select[1:i]
    
    tryCatch(
      {
        mses[i] <- cross_valid(preds, resp, 1, 2)
      }, 
      error = function(err) {
        print(paste('Error:', err))
      }
    )
  }
  
  # store selected best subset 
  forward_select <- data.frame(var_inc = 1:10,
                               mse = mses)
  var_subset[[resp]] <- full_model_select[1:forward_select$var_inc[which(
    forward_select$mse == min(forward_select$mse))]]
  
  print(paste(resp, 'done...'))
}
```


# Model Fitting

Tune the INGARCH model by looping through parameter values and calling the 
`cross_valid` function above:
```{r}
# define parameters
Pmax <- 2
Qmax <- 2

# define empty data frame for storing tuned values
tuned <- data.frame(resp = resp_names, 
                    pred = rep(NA, length(rep_names)),
                    p = rep(NA, length(rep_names)),
                    q = rep(NA, length(rep_names)))
                    
# establish start time of tuning process
start <- Sys.time()

# loop over response variables
for (resp in resp_names) {
  # define empty data frame to store MSE results
  tuning <- data.frame(p = rep(1:Pmax, each = Qmax),
                     q = rep(1:Qmax, times = Pmax),
                     mse = rep(NA, (Pmax * Qmax)))
  
  for (p in 1:Pmax) {
    for (q in 1:Qmax) {
      tuning$mse[which(tuning$p == p & tuning$q == q)] <- cross_valid(
        var_subset[[resp]], resp, p, q)
      
      print(paste('Done with p =', p, '& q=', q, 'in', Sys.time() - start, 
                  'minutes...'))
    }
  }
  
  tuned_p <- tuning$p[which(tuning$mse == min(tuning$mse))]
  tuned_q <- tuning$q[which(tuning$mse == min(tuning$mse))]
  tuned[which(tuned$resp == resp), ] <- c(resp, var_subset[[resp]], 
                                          tuned_p, tuned_q)
}

tuned
```

Collect fitted model coefficientestimates from one-to-many models into data 
frame:
```{r}
# define empty data frame to store coefficients
coefs <- data.frame(resp = resp_names,
                       beta.1 = rep(NA, 15),
                       beta.2 = rep(NA, 15),
                       beta.3 = rep(NA, 15),
                       beta.4 = rep(NA, 15),
                       beta.5 = rep(NA, 15),
                       a.740 = rep(NA, 15),
                       CA3.1 = rep(NA, 15),
                       CA3.2 = rep(NA, 15),
                       CA3.3 = rep(NA, 15),
                       CA3.4 = rep(NA, 15),
                       CA3.5 = rep(NA, 15),
                       CA3.6 = rep(NA, 15),
                       CA3.7 = rep(NA, 15),
                       CA3.8 = rep(NA, 15),
                       CA3.9 = rep(NA, 15),
                       CA3.10 = rep(NA, 15),
                       CA3.11 = rep(NA, 15),
                       CA3.12 = rep(NA, 15),
                       CA3.13 = rep(NA, 15),
                       CA3.14 = rep(NA, 15),
                       CA3.15 = rep(NA, 15),
                       CA3.16 = rep(NA, 15),
                       CA3.17 = rep(NA, 15),
                       CA3.18 = rep(NA, 15),
                       CA3.19 = rep(NA, 15),
                       CA3.20 = rep(NA, 15),
                       CA3.21 = rep(NA, 15))

# loop through response variables
for (region in resp_names) {
  dfs <- gen_df(1:10, 11, pred_names, resp)
  train_df <- dfs[['train_df']]
  remove(dfs)

  tryCatch(
    {
      model <- tsglm(as.ts(train_df[, 1]), xreg = as.matrix(train_df[, -1]), 
               model = list(past_obs = 1:5, past_mean = 740), 
               link = 'log', distr = 'poisson')
      temp <- summary(model)
      est <- temp$coefficients$Estimate[-1]
      coefs[which(coefs$resp == region), ] <- c(region, est)
      print(paste('Success:', region))
    },
    error=function(cond) {
      coefs[which(coefs$resp == region), ] <- c(region, 
                                                    rep(NA, ncol(train_df) + 5))
      print(paste('Error:', region))
    }
  )
}
```

Plot coefficient estimates from one-to-many models:
```{r}
coefs %>%
  pivot_longer(!resp, names_to = 'pred', values_to = 'est') %>%
  ggplot() +
  geom_tile(aes(factor(pred, levels = c('beta.1', 'beta.2', 'beta.3', 'beta.4', 
                                        'beta.5', 'a.740', pred_names)),
                factor(resp, levels = resp_names), 
                fill = as.numeric(est))) +
  scale_fill_continuous() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  labs(title='INGARCH Coefficient Estimates', 
       x='Predictor', y='Response', fill='')
```

Testing area for quick single model fitting: 
```{r}
p <- 5
q <- 10

# collect data frames
dfs <- gen_df(1:5, 10, pred_names, 'CA1.1')
train_df <- dfs[['train_df']]
valid_df <- dfs[['valid_df']]
valid_pred <- dfs[['valid_pred']]
remove(dfs)

# fit INGARCH model
start <- Sys.time()
model <- tsglm(as.ts(train_df[, 1]), xreg = as.matrix(train_df[, -1]), 
               model = list(past_obs = 1:p, past_mean = q), 
               link = 'log', distr = 'poisson')
print(Sys.time() - start)

summary(model)

# calculate MSE from predicted values
pred <- predict(model, n.ahead = nrow(valid_df), newxreg = valid_df)
mean((pred[['pred']] - valid_pred$CA1.1)^2, na.rm = TRUE)
```

