---
title: "Results"
author: "Adam Kiehl"
date: "4/7/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

Read in binned data:
```{r}
binned_df <- read.csv('../Data/binned_df.csv')

resp_names <- c('CA1.1', 'CA1.2', 'CA1.3', 'CA1.4', 'CA1.5', 'CA1.6', 'CA1.7', 
                'CA1.8', 'CA1.9', 'CA1.10', 'CA1.11', 'CA1.12', 'CA1.13', 
                'CA1.14', 'CA1.15')
pred_names <- c('CA3.1', 'CA3.2', 'CA3.3', 'CA3.4', 'CA3.5', 'CA3.6', 'CA3.7', 
                'CA3.8', 'CA3.9', 'CA3.10', 'CA3.11', 'CA3.12', 'CA3.13', 
                'CA3.14', 'CA3.15', 'CA3.16', 'CA3.17', 'CA3.18', 'CA3.19', 
                'CA3.20', 'CA3.21')

trls <- unique(binned_df$trial)
```

Exploratory view of data frame:
```{r}
summary(binned_df)
```

Plot spike trains for a single trial:
```{r}
binned_df %>%
  filter(trial == unique(binned_df$trial)[1]) %>%
  select(c(binstart, pred_names, resp_names)) %>%
  pivot_longer(c(all_of(pred_names), all_of(resp_names)), names_to = 'neuron', values_to = 'count') %>%
  ggplot() +
  geom_col(aes(binstart, count)) +
  facet_wrap(factor(neuron, levels = c(all_of(resp_names), all_of(pred_names))) ~ .) +
  theme_bw() +
  theme(axis.text.x = element_blank(), axis.text.y = element_text(size = 5)) +
  theme(axis.ticks.x = element_blank(), axis.ticks.y = element_line(size = .1)) +
  coord_cartesian(ylim = c(0, 30)) +
  scale_y_continuous(breaks = c(0, 10, 20, 30)) +
  labs(title = 'Single Trial Spike Trains', subtitle = 'Data binned into 0.5 second intervals', 
       x = 'Time [sec]', y = 'Bin Spike Count')
```

Calculate normalized cross-correlation matrix:
```{r}
norm_cc <- matrix(NA, nrow = length(c(pred_names, resp_names)), 
                  ncol = length(c(pred_names, resp_names))) %>% 
  as.data.frame() %>%
  mutate(var2 = c(pred_names, resp_names))
names(norm_cc) <- c(pred_names, resp_names, 'var2')
                     
for (var1 in c(pred_names, resp_names)) {
  for (var2 in c(pred_names, resp_names)) {
    if (var1 == var2) {
      break
    }
    
    temp <- ccf(binned_df[, var1], binned_df[, var2], plot = FALSE)
    norm_cc[which(norm_cc$var2 == var2), var1] <- max(temp$acf)
  }
}

var2_vec <- c()
for (i in 1:nrow(norm_cc)) {
  var2_vec <- c(var2_vec, rep(norm_cc$var2[i], (37 - i - 1)))
}

norm_cc %>%
  select(-var2) %>%
  pivot_longer(cols = c(pred_names, resp_names), 
               names_to = 'var1', values_to = 'corr') %>%
  na.omit(corr) %>%
  mutate(var2 = var2_vec) %>%
  ggplot() +
  geom_tile(aes(factor(var1, levels = c(pred_names, resp_names)), 
                factor(var2, levels = c(pred_names, resp_names)), 
                fill = corr)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = .65)) +
  labs(title = 'Normalized Cross Correlation Matrix', 
       subtitle = 'Max of lagged correlation values taken', 
       x = '', y = '', fill = 'Correlation\nCoefficient')
```

Collect test MSE results:
```{r}
test_mse_all <- data.frame(response = resp_names, 
                           GLAR_Pois_IR = readRDS('../Optimize/GLAR_Pois_IR/test_mses.RData')$mse,
                           GLAR_Pois_WR = readRDS('../Optimize/GLAR_Pois_WR/test_mses.RData')$mse,
                           GLAR_NBinom_IR = readRDS('../Optimize/GLAR_NBinom_IR/test_mses.RData')$mse,
                           GLAR_NBinom_WR = readRDS('../Optimize/GLAR_NBinom_WR/test_mses.RData')$mse,
                           INGARCH_Pois_IR = readRDS('../Optimize/INGARCH_Pois_IR/test_mses.RData')$mse,
                           INGARCH_Pois_WR = readRDS('../Optimize/INGARCH_Pois_WR/test_mses.RData')$mse,
                           INGARCH_NBinom_IR = readRDS('../Optimize/INGARCH_NBinom_IR/test_mses.RData')$mse,
                           INGARCH_NBinom_WR = readRDS('../Optimize/INGARCH_NBinom_WR/test_mses.RData')$mse,
                           GLARMA_IR = readRDS('../Optimize/GLARMA_Pois_IR/test_mses.RData')$mse,
                           GLARMA_WR = readRDS('../Optimize/GLARMA_Pois_WR/test_mses.RData')$mse)
test_mse_all

test_mse_all %>%
  pivot_longer(names(test_mse_all)[-1], names_to = 'model', values_to = 'MSE') %>%
  ggplot() +
  geom_tile(aes(factor(model, levels = names(test_mse_all)[-1]), 
                factor(response, levels = resp_names), fill = MSE)) +
  theme_bw() +
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank()) +
  theme(panel.grid = element_blank()) +
  coord_cartesian(ylim = c(-1, 15)) +
  annotate(geom = 'text', x = 1:10, y = 0, label = rep(c('IR', 'WR'), 5), size = 2) +
  annotate(geom = 'text', x = c(1.5, 3.5, 5.5, 7.5, 9.5), y = -.4, 
           label = c(rep(c('Pois', 'NBinom'), 2), 'Pois'), size = 3) +
  annotate(geom = 'text', x = c(2.5, 6.5, 9.5), y = -1, label = c('GLAR', 'INGARCH', 'GLARMA'), size = 3.5) +
  geom_segment(aes(x = 1.5, xend = 1.5, y = .5, yend = -.05), size = .2) +
  geom_segment(aes(x = 3.5, xend = 3.5, y = .5, yend = -.05), size = .2) +
  geom_segment(aes(x = 5.5, xend = 5.5, y = .5, yend = -.05), size = .2) +
  geom_segment(aes(x = 7.5, xend = 7.5, y = .5, yend = -.05), size = .2) +
  geom_segment(aes(x = 9.5, xend = 9.5, y = .5, yend = -.05), size = .2) +
  geom_segment(aes(x = 2.5, xend = 2.5, y = .5, yend = -.5), size = .2) +
  geom_segment(aes(x = 6.5, xend = 6.5, y = .5, yend = -.5), size = .2) +
  geom_segment(aes(x = 4.5, xend = 4.5, y = .5, yend = -1.5), size = .2) +
  geom_segment(aes(x = 8.5, xend = 8.5, y = .5, yend = -1.5), size = .2) +
  labs(title = 'Collected Test MSE Values', x = 'Model', y = 'Response Variable')
```

Inter-region GLAR poisson coefficients:
```{r}
readRDS('../Optimize/GLAR_Pois_IR/coefs.RData') %>%
  pivot_longer(!resp, names_to = 'pred', values_to = 'est') %>%
  mutate(pred = str_replace_all(pred, 'beta', 'Beta')) %>%
  mutate(pred = str_replace_all(pred, '_', ' ')) %>%
  mutate(pred = str_replace_all(pred, 'trial', 'Trial')) %>%
  mutate(pred = str_replace_all(pred, 'alpha', 'Alpha')) %>%
  ggplot() +
  geom_tile(aes(factor(pred, levels = c('Beta 1', 'Beta 2', 'Beta 3', 'Beta 4', 
                                        'Beta 5', 'Alpha', pred_names, 
                                        'Trial')),
                factor(resp, levels = resp_names), 
                fill = as.numeric(est))) +
  scale_fill_continuous() +
  scale_x_discrete(limits = c('Beta 1', 'Beta 2', 'Beta 3', 'Beta 4', 'Beta 5', pred_names, 'Trial')) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  labs(title='Inter-Region Poisson GLAR Coefficient Estimates', 
       x='Predictor', y='Response', fill='')
```

Within-region GLAR poisson coefficients:
```{r}
readRDS('../Optimize/GLAR_Pois_WR/coefs.RData') %>%
  pivot_longer(!resp, names_to = 'pred', values_to = 'est') %>%
  mutate(pred = str_replace_all(pred, 'beta', 'Beta')) %>%
  mutate(pred = str_replace_all(pred, '_', ' ')) %>%
  mutate(pred = str_replace_all(pred, 'trial', 'Trial')) %>%
  mutate(pred = str_replace_all(pred, 'alpha', 'Alpha')) %>%
  filter(pred != 'Alpha') %>%
  ggplot() +
  geom_tile(aes(factor(pred, levels = c('Beta 1', 'Beta 2', 'Beta 3', 'Beta 4', 
                                        'Beta 5', 'Alpha', resp_names, 
                                        'Trial')),
                factor(resp, levels = resp_names), 
                fill = as.numeric(est))) +
  scale_fill_continuous() +
  scale_x_discrete(limits = c('Beta 1', 'Beta 2', 'Beta 3', 'Beta 4', 'Beta 5', resp_names, 'Trial')) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  labs(title='Within-Region Poisson GLAR Coefficient Estimates', 
       x='Predictor', y='Response', fill='')
```

Inter-region GLAR negative binomial coefficients:
```{r}
readRDS('../Optimize/GLAR_NBinom_IR/coefs.RData') %>%
  pivot_longer(!resp, names_to = 'pred', values_to = 'est') %>%
  mutate(pred = str_replace_all(pred, 'beta', 'Beta')) %>%
  mutate(pred = str_replace_all(pred, '_', ' ')) %>%
  mutate(pred = str_replace_all(pred, 'trial', 'Trial')) %>%
  mutate(pred = str_replace_all(pred, 'alpha', 'Alpha')) %>%
  ggplot() +
  geom_tile(aes(factor(pred, levels = c('Beta 1', 'Beta 2', 'Beta 3', 'Beta 4', 
                                        'Beta 5', 'Alpha', pred_names, 
                                        'Trial')),
                factor(resp, levels = resp_names), 
                fill = as.numeric(est))) +
  scale_fill_continuous() +
  scale_x_discrete(limits = c('Beta 1', 'Beta 2', 'Beta 3', 'Beta 4', 'Beta 5', pred_names, 'Trial')) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  labs(title='Inter-Region Negative Binomial GLAR Coefficient Estimates', 
       x='Predictor', y='Response', fill='')
```

Within-region GLAR negative binomial coefficients:
```{r}
readRDS('../Optimize/GLAR_NBinom_WR/coefs.RData') %>%
  pivot_longer(!resp, names_to = 'pred', values_to = 'est') %>%
  mutate(pred = str_replace_all(pred, 'beta', 'Beta')) %>%
  mutate(pred = str_replace_all(pred, '_', ' ')) %>%
  mutate(pred = str_replace_all(pred, 'trial', 'Trial')) %>%
  mutate(pred = str_replace_all(pred, 'alpha', 'Alpha')) %>%
  filter(pred != 'Alpha') %>%
  ggplot() +
  geom_tile(aes(factor(pred, levels = c('Beta 1', 'Beta 2', 'Beta 3', 'Beta 4', 
                                        'Beta 5', 'Alpha', resp_names, 
                                        'Trial')),
                factor(resp, levels = resp_names), 
                fill = as.numeric(est))) +
  scale_fill_continuous() +
  scale_x_discrete(limits = c('Beta 1', 'Beta 2', 'Beta 3', 'Beta 4', 'Beta 5', resp_names, 'Trial')) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  labs(title='Within-Region Negative Binomial GLAR Coefficient Estimates', 
       x='Predictor', y='Response', fill='')
```

Inter-region INGARCH poisson coefficients:
```{r}
readRDS('../Optimize/INGARCH_Pois_IR/coefs.RData') %>%
  pivot_longer(!resp, names_to = 'pred', values_to = 'est') %>%
  mutate(pred = str_replace_all(pred, 'beta', 'Beta')) %>%
  mutate(pred = str_replace_all(pred, '_', ' ')) %>%
  mutate(pred = str_replace_all(pred, 'trial', 'Trial')) %>%
  mutate(pred = str_replace_all(pred, 'alpha', 'Alpha')) %>%
  ggplot() +
  geom_tile(aes(factor(pred, levels = c('Beta 1', 'Beta 2', 'Beta 3', 'Beta 4', 
                                        'Beta 5', 'Alpha', pred_names, 
                                        'Trial')),
                factor(resp, levels = resp_names), 
                fill = as.numeric(est))) +
  scale_fill_continuous() +
  scale_x_discrete(limits = c('Beta 1', 'Beta 2', 'Beta 3', 'Beta 4', 'Beta 5', 'Alpha', pred_names, 'Trial')) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  labs(title='Inter-Region Poisson INGARCH Coefficient Estimates', 
       x='Predictor', y='Response', fill='')
```

Within-region INGARCH poisson coefficients:
```{r}
readRDS('../Optimize/INGARCH_Pois_WR/coefs.RData') %>%
  pivot_longer(!resp, names_to = 'pred', values_to = 'est') %>%
  mutate(pred = str_replace_all(pred, 'beta', 'Beta')) %>%
  mutate(pred = str_replace_all(pred, '_', ' ')) %>%
  mutate(pred = str_replace_all(pred, 'trial', 'Trial')) %>%
  mutate(pred = str_replace_all(pred, 'alpha', 'Alpha')) %>%
  ggplot() +
  geom_tile(aes(factor(pred, levels = c('Beta 1', 'Beta 2', 'Beta 3', 'Beta 4', 
                                        'Beta 5', 'Alpha', resp_names, 
                                        'Trial')),
                factor(resp, levels = resp_names), 
                fill = as.numeric(est))) +
  scale_fill_continuous() +
  scale_x_discrete(limits = c('Beta 1', 'Beta 2', 'Beta 3', 'Beta 4', 'Beta 5', 'Alpha', resp_names, 'Trial')) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  labs(title='Within-Region Poisson INGARCH Coefficient Estimates', 
       x='Predictor', y='Response', fill='')
```

Inter-region INGARCH negative binomial coefficients:
```{r}
readRDS('../Optimize/INGARCH_NBinom_IR/coefs.RData') %>%
  pivot_longer(!resp, names_to = 'pred', values_to = 'est') %>%
  mutate(pred = str_replace_all(pred, 'beta', 'Beta')) %>%
  mutate(pred = str_replace_all(pred, '_', ' ')) %>%
  mutate(pred = str_replace_all(pred, 'trial', 'Trial')) %>%
  mutate(pred = str_replace_all(pred, 'alpha', 'Alpha')) %>%
  ggplot() +
  geom_tile(aes(factor(pred, levels = c('Beta 1', 'Beta 2', 'Beta 3', 'Beta 4', 
                                        'Beta 5', 'Alpha', pred_names, 
                                        'Trial')),
                factor(resp, levels = resp_names), 
                fill = as.numeric(est))) +
  scale_fill_continuous() +
  scale_x_discrete(limits = c('Beta 1', 'Beta 2', 'Beta 3', 'Beta 4', 'Beta 5', 'Alpha', pred_names, 'Trial')) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  labs(title='Inter-Region Negative Binomial INGARCH Coefficient Estimates', 
       x='Predictor', y='Response', fill='')
```

Within-region INGARCH negative binomial coefficients:
```{r}
readRDS('../Optimize/INGARCH_NBinom_WR/coefs.RData') %>%
  pivot_longer(!resp, names_to = 'pred', values_to = 'est') %>%
  mutate(pred = str_replace_all(pred, 'beta', 'Beta')) %>%
  mutate(pred = str_replace_all(pred, '_', ' ')) %>%
  mutate(pred = str_replace_all(pred, 'trial', 'Trial')) %>%
  mutate(pred = str_replace_all(pred, 'alpha', 'Alpha')) %>%
  ggplot() +
  geom_tile(aes(factor(pred, levels = c('Beta 1', 'Beta 2', 'Beta 3', 'Beta 4', 
                                        'Beta 5', 'Alpha', resp_names, 
                                        'Trial')),
                factor(resp, levels = resp_names), 
                fill = as.numeric(est))) +
  scale_fill_continuous() +
  scale_x_discrete(limits = c('Beta 1', 'Beta 2', 'Beta 3', 'Beta 4', 'Beta 5', 'Alpha', resp_names, 'Trial')) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  labs(title='Within-Region Negative Binomial INGARCH Coefficient Estimates', 
       x='Predictor', y='Response', fill='')
```

Inter-region GLARMA poisson coefficients:
```{r}
readRDS('../Optimize/GLARMA_Pois_IR/coefs.RData') %>%
  pivot_longer(!resp, names_to = 'pred', values_to = 'est') %>%
  mutate(pred = str_replace_all(pred, 'phi', 'Phi')) %>%
  mutate(pred = str_replace_all(pred, '_', ' ')) %>%
  mutate(pred = str_replace_all(pred, 'trial', 'Trial')) %>%
  mutate(pred = str_replace_all(pred, 'theta', 'Theta')) %>%
  ggplot() +
  geom_tile(aes(factor(pred, levels = c('Phi 1', 'Phi 2', 'Phi 3', 'Phi 4', 
                                        'Phi 5', 'Theta', pred_names, 
                                        'Trial')),
                factor(resp, levels = resp_names), 
                fill = as.numeric(est))) +
  scale_fill_continuous() +
  scale_x_discrete(limits = c('Phi 1', 'Phi 2', 'Phi 3', 'Phi 4', 'Phi 5', 'Theta', pred_names, 'Trial')) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  labs(title='Inter-Region Poisson GLARMA Coefficient Estimates', 
       x='Predictor', y='Response', fill='')
```

Within-region GLARMA poisson coefficients:
```{r}
readRDS('../Optimize/GLARMA_Pois_WR/coefs.RData') %>%
  pivot_longer(!resp, names_to = 'pred', values_to = 'est') %>%
  mutate(pred = str_replace_all(pred, 'phi', 'Phi')) %>%
  mutate(pred = str_replace_all(pred, '_', ' ')) %>%
  mutate(pred = str_replace_all(pred, 'trial', 'Trial')) %>%
  mutate(pred = str_replace_all(pred, 'theta', 'Theta')) %>%
  ggplot() +
  geom_tile(aes(factor(pred, levels = c('Phi 1', 'Phi 2', 'Phi 3', 'Phi 4', 
                                        'Phi 5', 'Theta', resp_names, 
                                        'Trial')),
                factor(resp, levels = resp_names), 
                fill = as.numeric(est))) +
  scale_fill_continuous() +
  scale_x_discrete(limits = c('Phi 1', 'Phi 2', 'Phi 3', 'Phi 4', 'Phi 5', 'Theta', resp_names, 'Trial')) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  labs(title='Within-Region Poisson GLARMA Coefficient Estimates', 
       x='Predictor', y='Response', fill='')
```
